Event-Related Potentials of the Semantically\\ Informed Perception of Unfamiliar Objects

### 2020-10-18 analysis/manuscript.Rmd
update plots, change condition names

How does our perception of an object change as soon as we discover what it is for? In this study, we presented participants with unfamiliar objects, preceded either by valid information about their function (leading to semantically informed perception) or by misleading information (leading to naive perception). Capitalizing on the high temporal resolution of event-related potentials (ERPs), we investigated how different stages in the processing hierarchy are modulated by this semantic understanding of objects before, during, and after it has been acquired.

@@

```{r exp1-preparation}
# Define function for reading and formatting behavioral data
read_behav <- function(txt_fname) {
  # Read log file
  dat <- suppressMessages(suppressWarnings(read_tsv(txt_fname)))
  # Rename columns for participant and item IDs
  dat %<>% rename(participant = VPNummer, item = StimID) %>%
    # Add a column for the part of the experiment
    mutate(part = case_when(
      Wdh %in% c(211, 231) ~ "I",
      Wdh %in% c(212, 232) ~ "II",
      Wdh %in% c(213, 233) ~ "III",
      Wdh == 234 ~ "IV"
    ) %>% as_factor()) %>%
    # Remove filler stimuli (i.e. well-known objects)
    filter(bek_unbek != "bekannt")
  # Assign stimuli to conditions based on keywords and button presses
  excl_known <- dat %>%
    filter(part == "I" & Tastencode %in% c(201, 251)) %>%
    pull(item)
  excl_insight <- dat %>%
    filter(part == "II" & grepl("richtig", Bed) & !Tastencode %in% c(201, 202, 251, 252)) %>%
    pull(item)
  excl_naive <- dat %>%
    filter(part == "II" & grepl("falsch", Bed) & !Tastencode %in% c(203, 204, 253, 254)) %>%
    pull(item)
  insight <- dat %>%
    filter(part == "II" & grepl("richtig", Bed) & Tastencode %in% c(201, 202, 251, 252)) %>%
    pull(item)
  naive <- dat %>%
    filter(part == "II" & grepl("falsch", Bed) & Tastencode %in% c(203, 204, 253, 254)) %>%
    pull(item)
  # Create a new condition for the experimental condition
  dat %<>% mutate(condition = case_when(
    item %in% excl_known ~ "Excl_known",
    item %in% excl_insight ~ "Excl_insight",
    item %in% excl_naive ~ "Excl_naive",
    item %in% insight ~ "Insight",
    item %in% naive ~ "Naive"
  ) %>%
    factor(levels = c("Insight", "Naive", "Excl_insight", "Excl_naive", "Excl_known")))
  # In Experiment 3, we need an additional column for position (upright vs. inverted)
  if ("richtig_inv" %in% dat$Bed) {
    dat %<>% mutate(position = as_factor(ifelse(grepl("inv", Bed), "Inverted", "Upright"))) %>%
      # Return only relevant columns
      select(part, position, condition, participant, item, RT)
  } else {
    # Return only relevant columns
    dat %<>% select(part, condition, participant, item, RT)
  }
  return(dat)
}

@@

# Define function to compute grand averages per condition
def compute_evokeds(epochs):
    evokeds = dict()
    evokeds_dat = dict()
    # Each part becomes a dictionary of conditions
    for pt in ["I", "II", "III"]:
        evokeds[pt] = dict()
        evokeds_dat[pt] = dict()
        # Each condition becomes a list of participants
        for cn in ["Insight", "Naive"]:
            evokeds[pt][cn] = list()
            evokeds_dat[pt][cn] = list()
            # For every participant we average trials separately for all parts and conditions
            for vp in epochs.metadata["participant"].unique():
                query = 'participant == "' + vp + '" and part == "' + pt + '" and condition == "' + cn + '"'
                evokeds[pt][cn].append(epochs[query].average())
            # Compute grand averages across participants for this part and condition
            evokeds[pt][cn] = mne.grand_average(evokeds[pt][cn])
            # Export only the actual data for R
            evokeds_dat[pt][cn] = evokeds[pt][cn].data
    return (evokeds, evokeds_dat)

@@

# # Plot evoked potentials
# import matplotlib.pyplot as plt
# mne.viz.plot_compare_evokeds(evokeds=evokeds_exp1['II'], picks='P8'); plt.show()
# evokeds_exp1['II']['Insight'].data = evokeds_exp1['II']['Insight'].data - evokeds_exp1['II']['Naive'].data
# evokeds_exp1['II']['Insight'].plot_topomap(times=[0.125, 0.175, 0.5], average=0.05, vmin=-1, vmax=1, cmap='viridis'); plt.show()
```

```{r exp1-analysis}
# Re-import behavioral data and re-factorize some columns
data_exp1 <- py$epochs_exp1$metadata %>% mutate(
  part = factor(part, levels = c("I", "II", "III")),
  condition = factor(condition, levels = c("Insight", "Naive", "Excl_insight", "Excl_naive", "Excl_known")),
  participant = factor(participant),
  item = factor(item)
)

@@

# Remove any trials excluded from conditions
data_exp1 %<>% filter(condition %in% c("Insight", "Naive")) %>% droplevels()

# Contrast coding for condition (insight-naive) and part (2-1, 3-1)
contrasts_condition <- t(cbind(c("Insight" = 1, "Naive" = -1)))
contrasts_part <- t(cbind(c("I" = -1, "II" = 1, "III" = 0), c("I" = 0, "II" = -1, "III" = 1)))
contrasts(data_exp1$condition) <- MASS::ginv(contrasts_condition)
contrasts(data_exp1$part) <- MASS::ginv(contrasts_part)

@@

# Define function to compute LMMs and follow-up contrasts for each component
compute_models <- function(dep, formula, data, control, specs) {
  require(lmerTest)
  require(buildmer)
  require(emmeans)
  build <- buildmer(
    control = control,
    REML = FALSE,
    buildmerControl = buildmerControl(
      formula = formula,
      data = data,
      direction = c("order", "backward"),
      elim = function(logp) exp(logp) >= .20,
      calc.anova = TRUE,
      ddf = "Satterthwaite",
      dep = dep
    )
  )
  emm_options(lmer.df = "Satterthwaite", lmerTest.limit = Inf)
  conts <- emmeans(build@model, specs = specs, infer = TRUE, data = data)$contrasts %>% as.data.frame()
  mod <- list("model" = build@model, "anova" = build@anova, "summary" = build@summary, "contrasts" = conts)
  return(mod)
}

@@

Participants for Experiment 1 were 24 German native speakers (13 female, 11 male) with a mean age of 24 years (range 18 to 31) and no history of psychological disorder or treatment. All participants were right-handed according to the Edinburgh inventory [@oldfield1971] and reported normal or corrected-to-normal vision. They gave written informed consent before starting the experiment and received a compensation of €8 per hour for participating. All experiments subsequently reported were carried out in accordance with the Declaration of Helsinki and approved by the local ethics committee.

@@

Stimuli for Experiments 1 and 2 consisted of 240 grayscale photographs of real-world objects, 120 of which were well-known everyday objects (e.g. a bicycle, a toothbrush), whereas the other 120 were rare objects presumed to be unfamiliar for the majority of participants (e.g. a galvanometer, an udu drum). The well-known objects were included as filler stimuli of no interest. A list of the unfamiliar objects can be found in Appendix A. All stimuli were presented on a light blue background with a size of 207 × 207 pixels on a 19-inch LCD monitor with a resolution of 1,280 × 1,024 pixels and a refresh rate of 75 Hz. At a standardized viewing distance of 90 cm, the images of the objects subtended approximately 3.9 degrees of participants' horizontal and vertical visual angle.

For each unfamiliar object, a noun--verb pair was selected describing the object's typical function or use in a way that could typically be related to its visual features and their configuration (e.g. current--measuring, pottery--drumming). In the *insight* condition, the presentation of each unfamiliar object was preceded by its respective description, whereas in the *naive* condition, it was preceded by a non-matching description belonging to one of the other objects. In both conditions, the noun and the verb were presented one above the other in black letters on a light blue background. For each participant, half of the unfamiliar objects were presented in the insight condition (i.e. with a matching description) and the other half were presented in the naive condition (i.e. with a non-matching description). All participants saw each unfamiliar object in only one of the two conditions and the assignment of objects to conditions was counterbalanced across participants. The experiment was programmed and displayed using Presentation® software (Neurobehavioral Systems, Inc., Berkeley, CA, www.neurobs.com).

Each experimental session consisted of three parts (Figure \@ref(fig:exp1-plot)A). In Part I, after written informed consent had been obtained and the EEG had been prepared, all 240 familiar and unfamiliar objects were presented once in random order and without verbal descriptions. Each trial consisted of a fixation cross presented in the middle of the screen for 0.5 s, followed by the presentation of the object until participants made a response or until a time out after 3 s. The inter-trial interval until the presentation of the next fixation cross was 0.5 s and participants took a self-timed break after each block of 60 objects. The task, which was kept the same throughout all parts and experiments, was to classify each object using one of four response alternatives: (a) "I know what this is or have a strong assumption," (b) "I have an assumption what this is," (c) "I have rather no assumption what this is," or (d) "I do not know what this is and have no assumption." Participants were asked to respond as quickly and as accurately as possible by pressing one out of four buttons with the index or middle finger of their left or right hand, respectively. The mapping of the rating scale to the four buttons (left to right or right to left) was counterbalanced across participants. Because we were interested in the acquisition of semantic knowledge about unfamiliar objects only, all objects for which the participant indicated knowing the object in Part I were excluded from further analysis (`r scales::percent(stims_exp1["Excl_known"]/120, accuracy = 0.1)` of rare objects per participant).

(ref:figure-1-caption) Procedure and results of Experiment 1. *(A)* In Part I, participants were presented with 120 unfamiliar objects and indicated whether they knew what kind of object they were viewing. In Part II, half of the objects were preceded by a matching description (in purple color for illustration), leading to semantic insight, and the other half with a non-matching description (in petrol color for illustration), leading to naive perception. In Part III, the same objects were presented again without the descriptions. *(B)* Bar plots and *(C)* ERP waveforms and scalp topographies separately for objects for which participants experienced semantic insight and naive perception in Parts I, II, and III. Insight was associated with significantly more negative amplitudes in the N1 component in Part II, significantly less negative amplitudes in the N400 component in Parts II and III, and significantly more positive amplitudes in the P1 component in Part III. Error bars show the standard error of the mean. Ampl. = amplitude.\newline\**p* \< .05. \*\**p* \< .01. \*\*\**p* \< .001.

```{r exp1-plot, include=TRUE, fig.height=12, fig.cap = "(ref:figure-1-caption)"}
# Manually extract significance levels from contrasts for plotting
stars_exp1 <- list(
  P1 = c(I = NA, II = NA, III = "*"),
  N1 = c(I = NA, II = "*", III = NA),
  N400 = c(I = NA, II = "***", III = "***")
)

@@

# Retrieve locations for the relevant electrodes from standard montage
fname_montage <- list.files(path = "renv/python", pattern = "easycap-M1.txt", recursive = TRUE, full.names = TRUE)[1]
montage <- py$evokeds_exp1$I$Insight$ch_names %>%
  as_tibble() %>%
  rename(electrode = value) %>%
  left_join(read_tsv(fname_montage) %>% rename(electrode = Site)) %>%
  mutate(
    x = eegUtils:::deg2rad(Theta) * cos(eegUtils:::deg2rad(Phi)),
    y = eegUtils:::deg2rad(Theta) * sin(eegUtils:::deg2rad(Phi))
  )

@@

  colour_insight <- viridisLite::viridis(1, begin = 0.1, end = 0.1)

@@

    geom_rect(aes(xmin = 18, xmax = 38, ymin = 13, ymax = 33), color = colour_insight, fill = "white") +
    geom_text(aes(x = 28, y = 23, label = "Potato\nmashing"), color = colour_insight, size = 10 / .pt,

@@

# Define function to plot ERP waveforms and topographies
plot_erps <- function(comps, evokeds, montage, stars) {
  pmap(comps, function(name, start, stop, roi) {
    # N400 gets a different scale than P1 and N1
    ymin <- ifelse(name == "N400", -3, -4)
    # Loop through parts
    parts <- map(c("I", "II", "III"), function(part) {
      # Create (conditions x time points) x electrodes tibble
      data_plot <- evokeds[[part]] %>%
        map(function(data) {
          data %>%
            t() %>%
            as_tibble(.name_repair = "unique") %>%
            set_colnames(montage$electrode) %>%
            mutate(.time = times, roi = rowMeans(select(., all_of(roi))))
        }) %>%
        bind_rows(.id = "condition")
      # Prepare waveform
      wave <- ggplot(data = data_plot, aes(x = .time, y = roi, color = condition)) +
        annotate("rect", xmin = start, xmax = stop, ymin = -Inf, ymax = ymin + 12, fill = "gray90")
      # Shade significant area
      star <- stars[[name]][[part]]
      if (!is.na(star)) {
        wave <- wave + geom_ribbon(
          data = data_plot %>% filter(.time >= start & .time <= stop) %>%
            group_by(.time) %>% summarise(ymin = min(roi), ymax = max(roi)),
          aes(x = .time, xmin = start, xmax = stop, ymin = ymin, ymax = ymax),
          inherit.aes = FALSE, fill = viridisLite::viridis(n = 1, begin = 1)
        )
      } #+
      # annotate("text", label = star, x = start + (stop - start)/2, y = -3.2, size = 7, family = "Helvetica")}
      # Add waves and styling
      wave <- wave +
        annotate("segment", x = -200, xend = 800, y = 0, yend = 0) +
        annotate("segment", x = 0, xend = 0, y = ymin, yend = ymin + 12) +
        annotate("segment", x = seq(-100, 700, 200), xend = seq(-100, 700, 200), y = -0.3, yend = 0) +
        annotate("segment", x = -12, xend = 0, y = seq(-2, 8, 4), yend = seq(-2, 8, 4)) +
        annotate("text", x = seq(-100, 700, 200), y = -0.9, label = seq(-100, 700, 200), size = 10 / .pt,
                 family = "Helvetica") +
        annotate("text", x = -20, y = seq(-2, 8, 4), label = seq(-2, 8, 4), size = 10 / .pt, family = "Helvetica",
                 hjust = 1) +
        annotate("text", x = 400, y = ymin + 0.8, label = "Time (ms)", size = 10 / .pt, family = "Helvetica",
                 lineheight = 0.9) +
        annotate("text",
          x = -100, y = 4, label = paste(name, "ampl.\n(µV)"), size = 10 / .pt,
          family = "Helvetica", angle = 90, lineheight = 0.9
        ) +
        geom_line() +
        scale_color_viridis_d(begin = 0.1, end = 0.5) +
        coord_cartesian(xlim = c(-200, 800), ylim = c(ymin, ymin + 14), expand = FALSE) +
        theme_void() +
        theme(legend.position = "none")
      # Plot topography
      topo <- data_plot %>%
        filter(.time >= start & .time <= stop) %>%
        group_by(condition) %>%
        summarise(across(montage$electrode, mean)) %>%
        pivot_longer(-condition) %>%
        pivot_wider(names_from = condition) %>%
        transmute(amplitude = Insight - Naive) %>%
        bind_cols(montage) %>%
        eegUtils::topoplot(limits = c(-1, 1), palette = "viridis", contour = FALSE, highlights = roi, scaling = 0.1) +
        theme(legend.position = "none")
      topo$layers[[3]]$aes_params$size <- topo$layers[[4]]$aes_params$size <- topo$layers[[5]]$aes_params$size <- 0.6
      # Combine waveform and topography
      wave + draw_plot(topo, width = 320, height = 9, x = 500, y = ymin + 6.4)
    })
    # Combine plots for the different parts
    plot_grid(parts[[1]], NULL, parts[[2]], NULL, parts[[3]], nrow = 1, rel_widths = c(10, 1, 10, 1, 10))
    # Combine plots for the different components
  }) %>% plot_grid(plotlist = ., nrow = 3)
}

>>>>

How does our perception of an object change as soon as we discover what it is for? In this study, we presented participants with unfamiliar objects, preceded in half of the cases by valid information about their function (leading to semantically informed perception) and in the other half of the cases by misleading information (leading to naive perception). Capitalizing on the high temporal resolution of event-related potentials (ERPs), we investigated if semantically informed versus naive perception led to differential effects at different stages in the processing hierarchy of visual objects.

Learning semantic knowledge about objects has an influence on how we process them [@abdelrahman2008; @gauthier2003; @gratton2009; @maier2014; @maier2019; @rossion2002; @rossion2004; @tanaka2001; @samaha2018; @weller2019]. Recently, evidence has accumulated that it does so not only at "higher" stages in the visual processing hierarchy, i.e. after "lower" stages of perceptual processing have been completed. It rather seems as though knowledge can also alter early processes, previously thought to be determined solely by the visual input itself, in a top-down manner. Such a top-down influence of semantic knowledge or expertise may manifest itself in behavioral changes such as improved visual matching [e.g.  @gauthier2003] and visual search performance [@maier2014; @holmes2012] when having learned that two objects share the same semantic category or verbal label. Furthermore, acquiring semantically meaningful knowledge about previously unfamiliar objects has been shown to increase their likelihood of gaining access to conuscious visual awareness under circumstances when attentional resources are limited [@weller2019]. Although these behavioral effects hint at an early locus of semantic knowledge effects in object processing, they in and of themselves cannot convincingly demonstrate that not just downstream, postperceptual steps are being affected [@firestone2016].

Stronger evidence comes from some of the above and other studies concurrently measuring event-related brain potentials (ERPs) while participants are performing different tasks on the objects about which they had previously acquired semantic knowledge. In their ERP study, [@abdelrahman2008] had participants first learn in-depth knowledge about 20 unfamiliar objects by listening to a verbal descripition of the function of each object while it was presented visually on the screen, whereas for 20 different objects they did not receive any relevant semantic information, instead listening to unrelated cooking recipes while the objects were presented. As in the present study, the assignment of unfamiliar objects to either of the two conditions was counterbalanced ascross participants as to prevent any subsequent effects of semantic knowledge from being confounded with visual differences between the objects. In a separate test session, event-related potentials were measured while participants performed three different tasks on those objects (intermixed with well known objects that had not been part of the training session). ERP amplitudes in response to objects for which in-depth versus minimal knowledge had been acquired differed in the time window of the N400 component (300--500 ms after stimulus onset), typically associated with semantic processing, but, interestingly, aslo in the time window of the visual P1 component (100--150 ms after stimulus onset), typcially associated with lower-level visual processing [[@abdelrahman2008]; Experiment 1]. This effect ocurred across three different tasks (none of which explicitly required accessing any of the aqcuired information) and was qualitatively similar to the ERPs observed for untrained but well-known objects. A modulation of the P1 component has also been replicated using similar sets of objects in the above mentioned studies showing effects of object knowledge in visual search and the visual awareness tasks [@maier2019; @weller2019]. It thus seems as though learning semantic information can influence steps of the processing hierarchy previously thought to be determined by the low-level visual input.

A different strategy to investaige the influence of semantic knowledge on perception has been taken in a recent ERP study by [@samaha2018]. Instead of using unfamiliar objects as stimuli, participants viewed images of familiar everyday objects that had been converted to two-tone "Mooney" images, making the object itself difficult to recognize. For half of the images, participants performed a semantic training where they were told which object to look for in each image. For the other half of images, they performed a perceptual (1-back) control task to equate the amount familiarity across all images. In a subsequent visual discrimination task, participants tended to respond faster and more accurate to the semantically trained images than to the perceptually trained ones [[@samaha2018]; Experiment 4]. Crucially, the semantic training also manifested itself in increased pre-stimulus alpha-band power and larger P1 amplitudes at left temporo-occipital electrodes, further coroborating an early influence of semantically informed as compared to naive perception of objects.

What all the previous studies have in common is that they investigated the effects of semantic knoweldge on object perception *after* an extensive learning phase [@abdelrahman2008; @gauthier2003; @maier2014; @maier2019; @rossion2002; @rossion2004; @samaha2018; @weller2019]. Participants usually performed a seperate training session where they encountered each object together with its respective label or description multiple times. The EEG was usually not measured (or at least not analyzed and reported) during this training phase, which sometimes took place on a separate day [@abdelrahman2008; @maier2014; @maier2019] or was even spread out across multiple days or weeks [@rossion2002; @rossion2004]. While designs with such substantive learning maximizes the chances of detecting even subtle top-down effects of the acquired knowledge, they necessarily leave at least three important conceptual questions unresolved. The first question is if we can detect any electrophysiological correlates of semantic insight, i.e. the critical presentation during which an understanding of a previously unfamiliar or meaningless objects is happening, instead of asking if this understanding leads to differential effects later in an orthogonal task. The second question is how much learning is actually necessary before reliable top-down effects of knowledge (e.g. on object-related P1 amplitudes) can be obtained: Does it actually take dozens of repitions per object or may a single exposure to the object together with the relevant semantic information be enough? The third and closely related question is related to one of the critiques sometimes raised against the nature of knowledge effects: Are they reflecting genuine top-down effects of semantic processes, altering perception online (while the object is being perceived), or do they merely reflect the (re-)activation of stored representations of the objects which have been created or altered throughout the learning process? By measuring the ERPs before, during, and after participants received semantic hints about the unfamiliar objects, we hope to provide tentative answers to all of these questions.

@@

```{r exp1-preparation}
# Define function for reading and formatting behavioral data
read_behav <- function(txt_fname) {
  # Read log file
  dat <- suppressMessages(suppressWarnings(read_tsv(txt_fname)))
  # Rename columns for participant and item IDs
  dat %<>% rename(participant = VPNummer, item = StimID) %>%
    # Add a column for the part of the experiment
    mutate(part = case_when(
      Wdh %in% c(211, 231) ~ "I",
      Wdh %in% c(212, 232) ~ "II",
      Wdh %in% c(213, 233) ~ "III",
      Wdh == 234 ~ "IV"
    ) %>% as_factor()) %>%
    # Remove filler stimuli (i.e. well-known objects)
    filter(bek_unbek != "bekannt")
  # Assign stimuli to conditions based on keywords and button presses
  excl_known <- dat %>%
    filter(part == "I" & Tastencode %in% c(201, 251)) %>%
    pull(item)
  excl_informed <- dat %>%
    filter(part == "II" & grepl("richtig", Bed) & !Tastencode %in% c(201, 202, 251, 252)) %>%
    pull(item)
  excl_naive <- dat %>%
    filter(part == "II" & grepl("falsch", Bed) & !Tastencode %in% c(203, 204, 253, 254)) %>%
    pull(item)
  informed <- dat %>%
    filter(part == "II" & grepl("richtig", Bed) & Tastencode %in% c(201, 202, 251, 252)) %>%
    pull(item)
  naive <- dat %>%
    filter(part == "II" & grepl("falsch", Bed) & Tastencode %in% c(203, 204, 253, 254)) %>%
    pull(item)
  # Create a new condition for the experimental condition
  dat %<>% mutate(condition = case_when(
    item %in% excl_known ~ "Excl_known",
    item %in% excl_informed ~ "Excl_informed",
    item %in% excl_naive ~ "Excl_naive",
    item %in% informed ~ "Informed",
    item %in% naive ~ "Naive"
  ) %>%
    factor(levels = c("Informed", "Naive", "Excl_informed", "Excl_naive", "Excl_known")))
  # In Experiment 3, we need an additional column for position (upright vs. inverted)
  if ("richtig_inv" %in% dat$Bed) {
    dat %<>% mutate(position = as_factor(ifelse(grepl("inv", Bed), "Inverted", "Upright"))) %>%
      # Return only relevant columns
      select(part, position, condition, participant, item, RT)
  } else {
    # Return only relevant columns
    dat %<>% select(part, condition, participant, item, RT)
  }
  return(dat)
}

@@

# Define function to compute grand averages per condition
def compute_evokeds(epochs):
    evokeds = dict()
    evokeds_dat = dict()
    # Each part becomes a dictionary of conditions
    for pt in ["I", "II", "III"]:
        evokeds[pt] = dict()
        evokeds_dat[pt] = dict()
        # Each condition becomes a list of participants
        for cn in ["Informed", "Naive"]:
            evokeds[pt][cn] = list()
            evokeds_dat[pt][cn] = list()
            # For every participant we average trials separately for all parts and conditions
            for vp in epochs.metadata["participant"].unique():
                query = 'participant == "' + vp + '" and part == "' + pt + '" and condition == "' + cn + '"'
                evokeds[pt][cn].append(epochs[query].average())
            # Compute grand averages across participants for this part and condition
            evokeds[pt][cn] = mne.grand_average(evokeds[pt][cn])
            # Export only the actual data for R
            evokeds_dat[pt][cn] = evokeds[pt][cn].data
    return (evokeds, evokeds_dat)

@@

# # Plot evoked potentials
# import matplotlib.pyplot as plt
# mne.viz.plot_compare_evokeds(evokeds=evokeds_exp1['II'], picks='P8'); plt.show()
# evokeds_exp1['II']['Informed'].data = evokeds_exp1['II']['Informed'].data - evokeds_exp1['II']['Naive'].data
# evokeds_exp1['II']['Informed'].plot_topomap(times=[0.125, 0.175, 0.5], average=0.05, vmin=-1, vmax=1, cmap='viridis'); plt.show()
```

```{r exp1-analysis}
# Re-import behavioral data and re-factorize some columns
data_exp1 <- py$epochs_exp1$metadata %>% mutate(
  part = factor(part, levels = c("I", "II", "III")),
  condition = factor(condition, levels = c("Informed", "Naive", "Excl_informed", "Excl_naive", "Excl_known")),
  participant = factor(participant),
  item = factor(item)
)

@@

# Remove any trials excluded from conditions
data_exp1 %<>% filter(condition %in% c("Informed", "Naive")) %>% droplevels()

# Contrast coding for condition (informed-naive) and part (2-1, 3-1)
contrasts_condition <- t(cbind(c("Informed" = 1, "Naive" = -1)))
contrasts_part <- t(cbind(c("I" = -1, "II" = 1, "III" = 0), c("I" = 0, "II" = -1, "III" = 1)))
contrasts(data_exp1$condition) <- MASS::ginv(contrasts_condition)
contrasts(data_exp1$part) <- MASS::ginv(contrasts_part)

@@

# Define function to compute LMMs and follow-up contrasts for each component
compute_models <- function(dep, formula, data, control, specs) {
  require(lmerTest)
  require(buildmer)
  require(emmeans)
  build <- buildmer(
    control = control,
    REML = FALSE,
    buildmerControl = buildmerControl(
      formula = formula,
      data = data,
      direction = c("backward", "backward"),
      elim = function(logp) exp(logp) >= .20,
      calc.anova = TRUE,
      ddf = "Satterthwaite",
      dep = dep
    )
  )
  emm_options(lmer.df = "Satterthwaite", lmerTest.limit = Inf)
  means <- emmeans(build@model, specs = specs, infer = TRUE, data = data)$emmeans %>% as.data.frame()
  conts <- emmeans(build@model, specs = specs, infer = TRUE, data = data)$contrasts %>% as.data.frame()
  mod <- list(model = build@model, anova = build@anova, summary = build@summary, means = means, contrasts = conts)
  return(mod)
}

@@

Participants for Experiment 1 were 24 German native speakers (13 female, 11 male) with a mean age of 24 years (range 18 to 31) and no history of psychological disorder or treatment. No a priori power analysis was carreid out and the sample size was chosen according to the default for EEG studies in our lab at the time of data collection. All participants were right-handed according to the Edinburgh inventory [@oldfield1971] and reported normal or corrected-to-normal vision. They gave written informed consent before starting the experiment and received a compensation of €8 per hour for participating. All experiments subsequently reported were carried out in accordance with the Declaration of Helsinki and approved by the local ethics committee.

@@

Stimuli for Experiments 1 and 2 consisted of 240 grayscale photographs of real-world objects, 120 of which were well-known everyday objects (e.g. a bicycle, a toothbrush), serving as filler stimuli of no interest, whereas the other 120 were rare objects presumed to be unfamiliar for the majority of participants (e.g. a galvanometer, an udu drum). A list of these unfamiliar objects can be found in Appendix A. All stimuli were presented on a light blue background with a size of 207 × 207 pixels on a 19-inch LCD monitor with a resolution of 1,280 × 1,024 pixels and a refresh rate of 75 Hz. At a standardized viewing distance of 90 cm, the images of the objects subtended approximately 3.9 degrees of participants' horizontal and vertical visual angle.

For each unfamiliar object, a pair of keywords---a noun and a verb---was selected describing the object's typical function or use in a way that could typically be related to its visual features and their configuration (e.g. current--measuring, pottery--drumming). As our central experimental manipulation, the presentation of each unfamiliar object was preceded by its correctly matching keywords for half of the objects, whereas the other half were preceded by non-matching keywords belonging to one of the other objects. The matching keywords were expected to induce semantically informed perception (i.e. participants suddenly understanding what kind of object they were viewing), whereas the non-matching keywords were expected hamper such an effect and keep the perception of the object semantically naive. All participants saw each unfamiliar object in only one of the two conditions and the assignment of objects to conditions was counterbalanced across participants. The experiment was programmed and displayed using Presentation® software (Neurobehavioral Systems, Inc., Berkeley, CA, www.neurobs.com).

Each experimental session consisted of three parts (Figure \@ref(fig:exp1-plot)A). In the *pre-insight* part (Part I), after written informed consent had been obtained and the EEG had been prepared, all 240 familiar and unfamiliar objects were presented once in random order and without any keywords. Each trial consisted of a fixation cross presented in the middle of the screen for 0.5 s, followed by the presentation of the object until participants made a response or until a time out after 3 s. The inter-trial interval until the presentation of the next fixation cross was 0.5 s and participants took a self-timed break after each block of 60 objects. The task, which was kept the same throughout all parts and experiments, was to classify each object using one of four response alternatives: (a) "I know what this is or have a strong assumption," (b) "I have an assumption what this is," (c) "I have rather no assumption what this is," or (d) "I do not know what this is and have no assumption." Participants were asked to respond as quickly and as accurately as possible by pressing one out of four buttons with the index or middle finger of their left or right hand, respectively. The mapping of the rating scale to the four buttons (left to right or right to left) was counterbalanced across participants.

(ref:figure-1-caption) Procedure and results of Experiment 1. *(A)* In the pre-insight part (Part I) participants were presented with 120 unfamiliar objects and indicated whether they knew what kind of object they were viewing. In the insight part (Part II), half of these objects were presented with matching keywords (in purple color for illustration), leading to semantically informed perception, and the other half with non-matching keywords (in petrol color for illustration), leading to naive perception. In the post-insight part (Part III), the same objects were presented again without the keywords. *(B)* ERP waveforms and scalp topographies separately for objects with semantically informed versus naive perception in Parts I, II, and III. Semantically informed perception was associated with significantly more negative amplitudes in the N1 component in Part II, significantly less negative amplitudes in the N400 component in Parts II and III, and significantly more positive amplitudes in the P1 component in Part III. Ampl. = amplitude.\newline\**p* \< .05. \*\**p* \< .01. \*\*\**p* \< .001.

```{r exp1-plot, include=TRUE, fig.height=11, fig.cap = "(ref:figure-1-caption)"}
# Manually extract significance levels from contrasts for plotting
stars_exp1 <- list(
  P1 = c(I = NA, II = NA, III = "*"),
  N1 = c(I = NA, II = "*", III = NA),
  N400 = c(I = NA, II = "***", III = "***")
)

@@

# Retrieve locations for the relevant electrodes from standard montage
fname_montage <- list.files(path = "renv/python", pattern = "easycap-M1.txt", recursive = TRUE, full.names = TRUE)[1]
montage <- py$evokeds_exp1$I$Informed$ch_names %>%
  as_tibble() %>%
  rename(electrode = value) %>%
  left_join(read_tsv(fname_montage) %>% rename(electrode = Site)) %>%
  mutate(
    x = eegUtils:::deg2rad(Theta) * cos(eegUtils:::deg2rad(Phi)),
    y = eegUtils:::deg2rad(Theta) * sin(eegUtils:::deg2rad(Phi))
  )

@@

  colour_informed <- viridisLite::viridis(1, begin = 0.1, end = 0.1)

@@

    geom_rect(aes(xmin = 18, xmax = 38, ymin = 13, ymax = 33), color = colour_informed, fill = "white") +
    geom_text(aes(x = 28, y = 23, label = "Potato\nmashing"), color = colour_informed, size = 10 / .pt,

@@

# Define function to plot ERP waveforms and topographies
plot_erps <- function(comps, evokeds, montage, stars) {
  suppressWarnings(suppressMessages(
    pmap(comps, function(name, start, stop, roi) {
      # N400 gets a different scale than P1 and N1
      ymin <- ifelse(name == "N400", -3, -4)
      # Loop through parts
      parts <- map(c("I", "II", "III"), function(part) {
        # Create (conditions x time points) x electrodes tibble
        data_plot <- evokeds[[part]] %>%
          map(function(data) {
            data %>%
              t() %>%
              as_tibble(.name_repair = "unique") %>%
              set_colnames(montage$electrode) %>%
              mutate(.time = times, roi = rowMeans(select(., all_of(roi))))
          }) %>%
          bind_rows(.id = "condition")
        # Shade background depending on whether the effect is significant or not
        star <- stars[[name]][[part]]
        if (is.na(star)) {
          shade <- annotate("rect", xmin = start, xmax = stop, ymin = ymin + 0.1, ymax = ymin + 11.9, color = "black", fill = NA)
        } else {
          shade <- annotate("rect", xmin = start, xmax = stop, ymin = ymin + 0.1, ymax = ymin + 11.9, color = "black", fill = "gray90")
        }
        xstar <- case_when(str_length(star) == 1 ~ start + 11, str_length(star) == 3 ~ start + 15)
        # # Prepare waveform
        # wave <- ggplot(data = data_plot, aes(x = .time, y = roi, color = condition)) +
        #   annotate("rect", xmin = start, xmax = stop, ymin = -Inf, ymax = ymin + 12, fill = "gray90")
        # # Shade significant area
        # star <- stars[[name]][[part]]
        # if (!is.na(star)) {
        #   wave <- wave + geom_ribbon(
        #     data = data_plot %>% filter(.time >= start & .time <= stop) %>%
        #       group_by(.time) %>% summarise(ymin = min(roi), ymax = max(roi)),
        #     aes(x = .time, xmin = start, xmax = stop, ymin = ymin, ymax = ymax),
        #     inherit.aes = FALSE, fill = viridisLite::viridis(n = 1, begin = 1)
        #   )
        # } +
        # annotate("text", label = star, x = start + (stop - start)/2, y = -3.2, size = 7, family = "Helvetica")}
        # Add waves and styling
